---
title: "course5progassignment2_PNM"
author: "PNM"
date: "January 11, 2018"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(utils)
library(RCurl)
```

# Course 5 - Programming Assignment 1 

## Read the data 

```{r}
setwd("C:\\Users\\mehtap\\Documents\\Coursera\\Course5")
filename <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"

if (url.exists(filename)) {
  download.file(filename,"stormdata.csv.bz2")
  
}

# unzip the file
unzip("stormdata.csv.bz2", exdir = "C:\\Users\\mehtap\\Documents\\Coursera\\Course5")
setwd("C:\\Users\\mehtap\\Documents\\Coursera\\Course5")

# Now read the CSV file and create the data frame

rawdata <- read.csv("stormdata.bz2",na.strings = "NA")
```

## Across United States, which type of events are most harmful with respect to population health?

What are some of the things I can do?
- facet wrap of the differnt type of events and histograms of injuries and fatalities
- show the distribution of the events and fatalities across United State states - some sort of map
```{r}
tempdata <- rawdata %>% 
  group_by(date) %>%
  summarise(sumsteps = sum(steps,na.rm = TRUE))
ggplot(tempdata) +geom_histogram(mapping = aes(x = sumsteps), na.rm = TRUE)

tempsub1 <- tempdata  %>% summarise(meansteps = mean(sumsteps , na.rm = T), mediansteps = median(sumsteps, na.rm = TRUE))

```

The mean steps are `r tempsub1$meansteps` and median steps are `r tempsub1$mediansteps`.


## Across United States, which types of events have the greatest economic consequences? 

```{r}
tempdata2 <- rawdata %>% group_by(as.factor(interval)) %>% summarise(avgsteps = mean(steps, na.rm = TRUE))

plot(x = tempdata2$`as.factor(interval)`, y= tempdata2$avgsteps, type = "l",
     main = "Average Steps taken by time interval",
     xlab = "Interval",
     ylab = "Step Count")
# find the location of the maximum point
maxindex <- which( tempdata2$avgsteps == max(tempdata2$avgsteps))

# this is the interval when the max steps are recorded, averaged across all days 
tempdata2$`as.factor(interval)`[maxindex]
```

