---
title: "course5progassignment2_PNM"
author: "PNM"
date: "January 11, 2018"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(utils)
library(RCurl)
```

# Course 5 - Programming Assignment 1 

## Read the data 

```{r}
setwd("C:\\Users\\mehtap\\Documents\\Coursera\\Course5")
filename <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"

if (url.exists(filename)) {
  download.file(filename,"stormdata.csv.bz2")
  
}

# unzip the file
unzip("stormdata.csv.bz2", exdir = "C:\\Users\\mehtap\\Documents\\Coursera\\Course5")
setwd("C:\\Users\\mehtap\\Documents\\Coursera\\Course5")

# Now read the CSV file and create the data frame

rawdata <- read.csv("stormdata.bz2",na.strings = "NA")
```

## What is mean total number of steps taken per day?
```{r}
tempdata <- rawdata %>% 
  group_by(date) %>%
  summarise(sumsteps = sum(steps,na.rm = TRUE))
ggplot(tempdata) +geom_histogram(mapping = aes(x = sumsteps), na.rm = TRUE)

tempsub1 <- tempdata  %>% summarise(meansteps = mean(sumsteps , na.rm = T), mediansteps = median(sumsteps, na.rm = TRUE))

```

The mean steps are `r tempsub1$meansteps` and median steps are `r tempsub1$mediansteps`.


## What is the average daily activity pattern?

```{r}
tempdata2 <- rawdata %>% group_by(as.factor(interval)) %>% summarise(avgsteps = mean(steps, na.rm = TRUE))

plot(x = tempdata2$`as.factor(interval)`, y= tempdata2$avgsteps, type = "l",
     main = "Average Steps taken by time interval",
     xlab = "Interval",
     ylab = "Step Count")
# find the location of the maximum point
maxindex <- which( tempdata2$avgsteps == max(tempdata2$avgsteps))

# this is the interval when the max steps are recorded, averaged across all days 
tempdata2$`as.factor(interval)`[maxindex]
```

## Imputing missing values 

```{r}
# Calculate and report the total number of missing values
TotalMissingValues <- sum(is.na(rawdata))
# Fill in the missing values 
# look up the interval of the missing value and fill in the mean from tempdata2
newdata <- rawdata
for (i in 1:nrow(newdata)) {
  if (is.na(newdata$steps[i])) {
    intval <- newdata$interval[i]
    fillinval <- tempdata2[as.factor(intval),2]
    newdata$steps[i] <- as.numeric(fillinval)
    
  }
}

# now create the histogram of the total steps taken per day 
tempdata3 <- newdata %>% 
  group_by(date) %>%
  summarise(sumsteps = sum(steps))
ggplot(tempdata3) +geom_histogram(mapping = aes(x = sumsteps))

# report out the mean and median values of the total steps 
tempsub2 <-tempdata3  %>% summarise(meansteps = mean(sumsteps), mediansteps = median(sumsteps))


```

The mean steps are `r tempsub2$meansteps` and median steps are `r tempsub2$mediansteps`.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# create a factor variable indicating if the date is a weekday or weekend 
weekdayvec <- weekdays(as.Date(newdata$date), abbr = TRUE)

newdata$weekend <- ifelse(weekdays(as.Date(newdata$date),abbreviate = TRUE) %in% c("Sat", "Sun"), "weekend", "weekday")

# now make a panel plot

tempdata4 <- newdata %>% group_by(Interval = as.factor(interval), Weekend = as.factor(weekend)) %>% summarise(avgsteps = mean(steps, na.rm = TRUE))

tempdata5 <- tempdata4 %>% filter( Weekend == "weekday")
tempdata6 <- tempdata4 %>% filter(Weekend == "weekend")
par(mfrow = c(2,1))
plot(tempdata5$Interval, tempdata5$avgsteps, type = "l", main = "Weekday")
plot(tempdata6$Interval, tempdata6$avgsteps, type = "l", main = "Weekend")
```
